An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-08-01 EconGrowth2.Rnw
Added resample scatter plots for (le)k and (ek)l nestings for CES.

<<Historical_Lattice_Graph, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="Historical data: indexed GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$).">>=
createHistoricalLatticeGraph()
@

@@

CES:

($kl$)$e$: energy is unnec ($\delta$ is 1); also ($el$)$k$ energy is unnec. 
($\delta_1$ is 0);

@@

$\sigma$:
capital and energy are complements
labour and energy are subsitutable

@@

CES:

($kl$)$e$: energy is unnec ($\delta =1$);

@@

$\sigma$: energy is not substitutable with capital or labour, totally complementary

@@

CES:

($kl$)$e$; ($le$)$k$; ($ek$)$l$: no indications, all are important.

@@

$\sigma$: labour and energy are complements, capital and energy complements

@@

CES:

($kl$)$e$: energy is unnec. ($\delta = 1$);

@@

$\sigma$: capital and energy is complementary, but labour and energy more substitutable

@@

CES:

($kl$)$e$: labour is unnec. ($\delta_1 = 1$);

@@

$\sigma$: capital and energy is complementary, labour and energy is substitutable

@@

CES: no indications, but weak signal that labour is unnec in ($ek$)$l$ ($\delta = 0.93$)

@@

$\sigma$: labour and energy are complements, capital and energy are substitutes

@@

CES:

($kl$)$e$: energy is unnec ($\delta=1$);

@@

$\sigma$: labour and energy still complements, 
but not as strong capital and energy complements

@@

CES:

($kl$)$e$: labour is unnec ($\delta_1=0.97$);

@@

$\sigma$: labour and energy complements, capital and energy perfect substitutes

@@

CES:

($kl$)$e$: labour is unnec. ($\delta_1=1$); labour is unnec

@@

$\sigma$: labour and energy complements, not as strong. 
Capital and energy perfect substitutes.

>>>>

<<Historical_Lattice_Graph, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="Historical data: indexed GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$).", eval=FALSE>>=
createHistoricalLatticeGraph()
@

@@

CES with $q$:

($kl$): $\delta_1 = 0.3, \sigma_1 = 0.8$. $k$ is less necessary. Complements, but close to substitutes.
signal of substitutability in the spread.

($kl$)$e$: energy is unnec ($\delta = 1$). $\delta_1 = 0.45$: $k$ and $l$ are about equally necessary.
$\sigma_1 = 0.8$: $k$ and $l$ are slight complements. Spread on $0.25 < \sigma_1 < \infty$, signalling 
possibilities of substitution.

($el$)$k$: energy is unnec. ($\delta_1$ is 0);

@@

CES with $u$:

($kl$)$e$: $\delta = 1$, signalling $e$ unnecessary. $\sigma = 0.8$, indicating clost to substitutability,
but still complementary. 
$\delta_1 = 0.25$, with large spread. not strong for $k$.
$\sigma_1 \rightarrow \infty$, but with large spread from 0.2 upward. Signals complementarity
between $k$ and $l$.

@@

CES with $q$:

($kl$): $\delta_1 = 0.7, \sigma_1 = 0.35$. $kl$ complementary, $k$ is more important, because $\delta_1 > 0.5$. Large spread on $\delta_1$. Complementary signal is consistent, however.

($kl$)$e$: energy is unnec ($\delta = 1$). $\delta_1 = 0.25$, $k$ is less necessary. 
$\sigma_1 \rightarrow \infty$, giving strong $kl$ substitutability signal.

@@

CES with $u$:

($kl$)$e$: $\delta = 1$, $u$ is unnecessary. $\sigma = 0.8$, approache complementarity between $kl$ and $e$.
$\delta_1 = 0.1$. $\sigma_1 \rightarrow \infty$, indicating substitutability between $k$ and $l$.

@@

CES with $q$:

($kl$): $\delta_1 = 0.7, \sigma_1 = 0.25$. Very little scatter. Both are well-known. 
$l$ is close to unnecessary. $k$ and $l$ are complements.

($kl$)$e$: no indications, all are important. $\delta = 0.6$, both $kl$ and $e$ are necessary.
$0.5 < \delta < 1$.
$\sigma=0.1$ signalling complementarity betwen $kl$ and $e$.
$\delta_1 = 0.8$ signalling $l$ is almost unnecessary. $\sigma_1 = 0.1$, signalling complementarity 
between $k$ and $l$.

($le$)$k$: no indications, all are important.

($ek$)$l$: no indications, all are important.

@@

CES with $u$:

($kl$)$e$: $\delta = 0.6, \sigma = 0.4$. Complementarity between $kl$ and $e$.
$\delta_1 = 1$, indicating $l$ unnecessary. $\sigma_1 = 0.1$, indicating complementarity
between $k$ and $l$. Need only one of $k$ or $l$ in your equation, but they are
complemetary! You need one to have the other.

@@

CES with $q$:

($kl$): $\delta_1 = 0.3$, so $k$ is getting close to unnecessary. $\sigma_1 = 1.25$.
$k$ and $l$ are substitutes. Large spread in $\delta_1$ from 0.1 to 1.0. Large spread in
$\sigma_1$, but most points have $\sigma_1 > 1$.

($kl$)$e$: $\delta = 0.8$. however, strong signal at $0.8 < \delta < 1$, 
signalling that energy is less necessary. $0.5 < \delta < 1$. $\sigma = 0.5$, 
signalling complementarity between $kl$ and $e$. 
$\delta_1 = 1$, signalling that $l$ is unnecessary. $\sigma_1 = 0.1$, signalling
that $k$ and $l$ are complementary. Both $\delta_1$ and $\sigma_1$ have large spreads, 
so our confidence in these results is low.

@@

@@

CES with $q$:

($kl$): $\delta_1 = 0.35, \sigma_1 = 0.8$. $k$ is relatively less necessary. $k$ and $l$ 
are complements but relatively close to being substitutes. Spread is tight between 
$0.25 < \delta_1 < 0.5$ and $\sigma_1$ is consistently complementary 
between $0.1 < \sigma_1 < 0.8$.

($kl$)$e$: labour is unnec. ($\delta_1 = 1$). $\delta = 0.25$, signalling complemtarity 
between $kl$ and $e$, but large spread from 0 to 1. $\sigma$ and $\sigma_1$ are both 0, 
indicating perfect complements betwen $kl$ and $e$ and between $k$ and $l$.

@@

@@

CES with $q$:

($kl$): $\delta_1 = 0.85$. $l$ is unnecessary. $\sigma_1 = 0.1$ indicating strong complementarity between 
$k$ and $l$. Spread of $0.15 < \delta_1 < 1$. $0 < \sigma_1 < 0.8$, signalling complementarity.

($kl$)$e$: $\delta = 0.6$, $e$ is less necessary (weak signal). Big spread on $\delta$ between 0 and 0.8.
$\sigma = 0.1$ indicating complementarity between $kl$ and $e$. $\delta_1 = 0.8$ 
(with large spread between 0.3 and 1), labor is less necessary.
$\sigma_1 = 0.1$: very complementary between $k$ and $l$.

($le$)$k$:

($ek$)$l$: weak signal that labour is unnec ($\delta = 0.93$)

@@

@@

CES with $q$:

($kl$): $\delta_1 = 0.3$. Spread $0 < \delta_1 < 0.7$. $0 < \sigma_1 < 0.8$. $\sigma_1 = 0$. 
Strong complementarity. $k$ is less necessary compared to $l$.

($kl$)$e$: energy is unnec ($\delta=1$). $\sigma = 0.75$ indicating complementarity between $kl$ and $e$, approaching substitutibility.
$\delta_1 = 0.3, \sigma_1 = 0$ indicating strong perfect complementarity between $k$ and $l$.

@@

@@

CES with $q$:

($kl$): $\delta_1 = 0.8$ $0.3 < \delta_1 < 1$ signalling labor is less necessary.
$\sigma_1 = 0.1$ $0 < \sigma_1 < 0.8$, signalling complementary $k$ and $l$.

($kl$)$e$: labour is unnec ($\delta_1=0.97$). $\delta = 0.45$, so $kl$ and $e$ both necessary, but 
$\sigma \rightarrow \infty$, so $kl$ and $e$ has perfect substitutibility.
$\sigma_1 = 0.1$, so $k$ and $l$ are perfect complements. Spread on $\delta$ is large (0.2 to 1).
So, we are unsure about the value of $\delta$.

@@

@@

CES with $q$:

($kl$): $\delta_1 = 1$: $l$ is unnecessary. $0.5 < \delta_1 < 1$. $\sigma_1 = 0.8$ without any 
noticeable spread indicating complemetary.

($kl$)$e$: labour is unnec. ($\delta_1=1$). $\delta = 0.6$, $kl$ and $e$ are both necessary. 
$\sigma \rightarrow \infty$, signalling substitutability between $kl$ and $e$. $\sigma_1 = 0.8$ with very
large spread between 0 and 2. Approaching substitutability between $k$ and $l$ and perfect substitutibility
between $kl$ and $e$. Very little certainty here.

@@
